\section{The Field of Real Numbers}\label{secton_1.3}

\begin{theorem}\label{theorem_1.3.1}
  There exists an ordered field $\R$ with the least upper bound property, such
  that $\Q \subseteq \R$.
\end{theorem}

\begin{definition}
  We call the field $\R$ the \textbf{field of real numbers},and we call the
  elements of $\R$ \textbf{real numbers}.
\end{definition}

\begin{definition}
  Let $S$ be an ordered field, and let  $E \subseteq S$. We say that  $E$ is
  \textbf{dense} in $S$, if for all $r,s \in S$, with $r<s$, there is an $\alpha
  \in E$ such that $r<\alpha<s$.
\end{definition}

\begin{theorem}[The Archimedean Principle]\label{theorem_1.3.2}
  If $x,y \in \R$, and  $x>0$, then there is an  $n \in \Z^+$ such that
  $nx>y$.
\end{theorem}
\begin{proof}
  Let $A=\{nx: n \in \Z^+\}$, and suppose that  $nx \leq y$. Then  $y$ is an
  upper bound of $A$, abd since  $A$ is non-empty,  $\alpha=\sup{A} \in \R$,
  since $x>0$, we have $\alpha-x<\alpha$, so  $\alpha-x$ is not an upper bound of
  $A$. Hence  $\alpha-x<mx$ for some $m \in \Z^+$. Then  $\alpha<(1-m)x \in A$,
  contradicting that  $\alpha$ is an upper bound of  $A$.
\end{proof}

\begin{theorem}[The density of $\Q$ in  $\R$]\label{theorem_1.3.3}
  $\Q$ is dense in  $\R$.
\end{theorem}
\begin{proof}
  Let $x<y$ be real numbers, then  $y-x>0$, so by the Archimedean principle,
  there is an $n \in \Z^+$ fir which $n(y-x)>1$. By the Archimedean principle
  again, we have $m_1,m_2 \in \Z^+$ for which $m_1>nx$ and $m_2>-nx$, thus
  $-m_2<nx<m_1$, and we also have that there is an $m \in \Z^+$ for which
  $-m_2<m<m_1$, and $m-1 \leq nx<m$. Thus combining inequalities, we get
  $nx<m<ny$, thus  $x<\frac{m}{n}<y$.
\end{proof}

\begin{theorem}\label{theorem_1.3.4}
  For every real number $x \geq 0$, and for every  $n \in \Z^+$, there is one,
  and only one nonnegative real number  $y$ for which  $y^n=x$.
\end{theorem}
\begin{proof}
  Let $y \geq 0$ be a real number; then $y^n \geq 0$, so there is at most one
  such $y$ for which $y^n=x$. Now let  $E=\{t:\R: t^n<x\}$, choosing
  $t=\frac{x}{1+x}$, we see that $0 \leq t<1$, hence  $t^n<t<x$, so $E$ is
  non-empty. Now if $1+x<t$, then $t^n \geq x$, so $t \not\in E$, and  $E$ has
  $1+x$ as an upper bound. Therefore, $\alpha=\sup{E} \in \R$ exists.

  Now suppose that  $y^n<x$, choose  $0 \leq h<1$ such that
  $h<\frac{x-y^n}{n(y+1)^{n-1}}$, then
  $(y+h)^n-y^n<hn(y+h)^{n-1}<hn(y+1)^{n-1}<x-y^n$, thus  $(y+h)^n<x$, so  $y+h
  \in E$, contraditing that  $y$ is an upper bound. On the other hand, if
  $y^n>x$, choosing $k=\frac{y^n-x}{ny^{n-1}}$, then $0 \leq k<y$, and letting
  $t \geq y-k$, we get that $y^n-t^n \leq y^n+(y-k)^n<kny_{n-1}=y^n-x^n$, so
  $t^n \geq x$, making  $y-k$ an uppeerbound of  $E$, which contradicts
  $y=\sup{E}$.
\end{proof}

\begin{definition}
  Let $x \geq 0$ be a real number. We call the real number $y$ for which $y^n=x$
  for some $n \in \Z^+$ the \textbf{$n$-th root} of $x$. We denote $y$ as
  $\sqrt[n]{x}$, or as  $x^{\frac{1}{n}}$.
\end{definition}

\begin{corollary}
  If $a,b \in \R$, with  $a,b>0$, and  $n \in \Z^+$, then
  $\sqrt[n]{ab}=\sqrt[n]{a}\sqrt[n]{b}$.
\end{corollary}
\begin{proof}
  Let $\alpha=\sqrt[n]{a}$, and  $\beta=\sqrt[n]{b}$. Then  $\alpha^n=a$, and
  $\beta^n=b$, so $ab=\alpha^n\beta^n=(l\alpha\beta)^n$, we are done.
\end{proof}

\begin{definition}
  We define the \textbf{extended real number system} to be the field $\R$,
  together with elements $\infty$, and  $-\infty$, called  \textbf{positive
  infinity} and \textbf{negative infinity}, such that $-\infty<x<\infty$ for
  all $x \in \R$. We call elements of the extended real numbers \textbf{infinite},
  and every other element not in the extended real numbers \textbf{finite}. We
  denote the extended real numbers by $\R^\infty$.
\end{definition}

\begin{lemma}\label{theorem_1.3.5}
  $\infty$ is an upper bound for every subset  $E$, of  $\R$, and  $-\infty$ is
  a lower bound for every subset  $E$ of  $\R$. Moreover, if $E$ is not bounded
  above in $\R$, then  $\sup{E}=\infty$, and if  $E$ is not bounded below,
  then  $\inf{E}=-\infty$.
\end{lemma}
\begin{proof}
  This comes directly from definition. Since $-\infty<x<\infty$ for every
  element $x \in E$, $\infty$ and $-\infty$ are upper and lower bounds for $E$
  respectively.

  Now suppose that $E$ is not bounded above in $\R$, and let $s=\sup{E}$.
  Suppose further that $s \neq \infty$. If $\infty<s$, then $x<s$ for every $x
  \in \R$, so that $s=\infty$, which is impossible, so  $s<\infty$. But then  $s
  \in \R$, and $x \leq s$ which makes $E$ bounded above, a contradiction!
  Therefore,  $s=\infty$. By similar reasoning, we get $\inf{E}=-\infty$.
\end{proof}

\begin{remark}
  We make the following assumptions for extended real numbers:
  \begin{enumerate}
    \item[(1)] If $x \in \R$, then  $x+\infty=\infty$, $x-\infty=-\infty$,
      and $\frac{x}{\infty}=\frac{x}{-\infty}=0$.

    \item[(2)] If $x>0$, then  $x(\infty)=\infty$ and  $x(-\infty)=-\infty$.

    \item[(3)] If  $x<0$, then  $x(\infty)=-\infty$ and  $x(-\infty)=\infty$.
  \end{enumerate}
\end{remark}
