\section{Definitions and Examples}
\label{section_1.1}

\begin{definition}
  Let $G$ be set, we define a  \textbf{binary operation}, $\ast$, on  $G$ to
  be a map  $\ast:G \times G \rightarrow G$ that takes $(a,b) \rightarrow a
  \ast b$. We say that a binary operation $\ast$ is \textbf{associative} if
  for any $a,b,c \in G$,  $(a \ast b) \ast c=a \ast (b \ast c)$. We say that a
  binary operation, $\ast$ is \textbf{commutative} if for any $a,b \in G$,
  $a \ast b=b \ast a$
\end{definition}
We also write $ab$ instead of  $a \ast b$ for convinience, and when
context is clear.

\begin{example}\label{example_1.1}
  \begin{enumerate}
    \item[(1)] The usual addition, $+$ is an associative and commutative binary
      operation on the sets $\Z$,  $\Q$, and  $\R$ of integers, rationals,
      and real numbers. The addition of complex numbers, $+$ is an
      associative and commutative binary operation on the complex numbers
      $\C$

    \item[(2)] The usual multiplication $\cdot$ is an associative and
      commutative binary operation on  $\Z^*$,  $\Q^*$ and  $\R^*$.
      Complex multiplication on  $\C^*$ is also an associative,
      commutative binary operation. Note we define
      $F^*=\com{{F}}{\{0\}}$, where $F=\Z,\Q,\R,\C$.

    \item[(3)] The usual subtration, is a noncomutative binary operation on
      $\Z$,  $\Q$, and  $\R$, complex subtraction is noncommutative on
      $\C$. The map $a \rightarrow -a$ is not binary.

    \item[(4)] The usual subtraction is not a binary operation on $\Z^+$,
      $\Q^+$, and  $\R^+$, notice that if  $a<b$,  $a-b \notin F$ where
      $F=\Z,\Q,\R$.

    \item[(5)] The cross prodcuct, $\times$ on two vectors in real $3$-space
      is a nonassociative binary operation on $\R^3$.

    \item[(6)] The operation $+_n$ of addition modulo $n$ is a binary
      operation on the set of integers modulo $n$, $\faktor{\Z}{n\Z}$.
      Notice that for $k, l, m \in \Z$, that $a \mod{n}$ is of the form
      $a+kn$, thus $b \mod{n}$ and $c \mod{n}$, have the forms $b+ln$ and
      $c+mn$. Thus
      \begin{align*}
        ((a+kn)+(b+ln))+(c+mn) &= (a+b)+(l+k)n+(c+mn) \\
                            &= a+b+c+(l+k+m)n \\
                            &=(a+kn)+(b+c)+(l+m)n \\
                            &=(a+kn)+((b+ln)+(c+mn)) \\
      \end{align*}
      This implies that $(a+b)+c \mod{n}=a(b+c) \mod{n}$; additionally,
      \begin{align*}
        (a+kn)+(b+ln) &= (a+b)+(k+l)n \\
                      &=(b+a)+(l+k)n  \\
                      &=(b+ln)+(a+kn) \\
      \end{align*}
      so $a+b \mod{n} = b+a \mod{n}$. That is $+_n$ is associative and
      commutative. We abbreviate addition modulo $n$ and write $+$ instead
      of  $+_n$.

    \item[(7)] Multiplictation moduleo $n$ is a binary operation on
      $\faktor{\Z}{n\Z}$ which is associative and commutative.

    \item[(8)] The operation of function composition $\circ$ is a binary
      operation on any set of mappings. We have that for mappings $f$,
      $g$, and  $h$ that  $(f \circ g) \circ h=f \circ (g \circ h)$,
      making $\circ$ associative; but  $f \circ g \neq g \circ f$, making
      $\circ$ noncommutative.
  \end{enumerate}
\end{example}

\begin{definition}
  Let $G$ be a set, and  $H \subseteq G$, and let  $\ast$ be a binary
  operation on  $G$. We say that  $H$ is \textbf{closed} under $\ast$ if
  $\ast|_H$ is a binary operation on  $H$.
\end{definition}

\begin{definition}
  Let $G$ be a nonempty set, and let  $\ast$ be a binary operation on $G$. We
  call the pair  $(G,\ast)$ a \textbf{group} if:
  \begin{enumerate}
    \item[(0)] For every $a,b \in G$,  $a \ast b \in G$. That is  $G$ is closed
      under $\ast$.
    \item[(1)] $(a \ast b) \ast c=a \ast (b \astc)$ for all $a,b,c \in G$,
      i.e. $\ast$ is associative.

    \item[(2)] There exists an element $e \in G$ called the
      \textbf{identity} element such that  $a \ast e=e \ast a=a$ for all
      $a \in G$

    \item[(3)] For each $a \in G$, there is an element  $b \in G$, called
      the \textbf{inverse} of $a$ such that  $a \ast b=b \asta=e$, where
      $e$ is the identity element of $G$.
  \end{enumerate}
\end{definition}
\begin{remark}
  We make note that property $(0)$ of this definition is impled by stating
  $\ast$ as a binary operation on  $G$, we however list it, becuase when
  verifying a given set is a group, we usually want to check for closure.
\end{remark}
\begin{remark}
  Instead of stating $(G,\ast)$ as a group, we will often just say that $G$ is
  a group under  $\ast$, or simply,  $G$ is a group.
\end{remark}

\begin{example}\label{example_1.2}
  \begin{enumerate}
    \item[(1)] The set $G=\{e\}$ of one element forms a group under any
      operation, and is called the \textbf{trivial} group. We write
      $G=\langle e \rangle$. We also observe that there is no group $G$ which
      is empty (that is $\0$ is not a group). Indeed, suppose we removed
      the restriction to be nonempty in the definition of a group. We see
      that if $G$ is empty, then  $G$ cannot be a group, since it is
      not closed, trivially; furthurmore, there is no identity, nor inverse
      to each element. Therefore the minimum number of elements a group can
      have is $1$. This makes the trivial group $\langle e \rangle$
      the smallest possible group.

    \item[(2)] The sets $\Z$,  $\R$, and  $\Q$ are all groups under the
      usual addition. Here $0$ is the identity, and  $-a$ is the inverse
      of  $a$.  $\C$ is a group under complex addition with $0=0+i0$ the
      identity and $-a-ib$ the inverse of $a+ib$.

    \item[(3)] $\Q$ and  $\R$ are groups under the usual multiplication,
      with identity  $1$ and inverse  $\inv{a}=\frac{1}{a}$. $\Z$ is not a
      group under this operation, as  $\frac{1}{a} \notin \Z$ whenever $a
      \in \Z$.  $\C$ is a group under complex multiplication with identity
      $1=1+i0$ and inverse $\frac{a}{a^2+b^2}-i\frac{b}{a^2+b^2}$ for
      $a+ib$.

    \item[(4)] Consider the set of integers modulo $n$, $\faktor{\Z}{n\Z}$
      under addition modulo $n$, $+$. Since  $+$ is a binary operation on
      $\faktor{Z}{n\Z}$, closure is guaranteed. We also see that
      associativity hold. Now, notcie that $n \equiv 0 \mod{n}$, by
      definition, so $a+n=n+a \equiv 0+a \mod{n} \equiv a+0 \mod{n}
      \equiv a$. Moreover, $n-a \equiv 0-a \equiv -a \mod{n}$, and
      $(n-a)+a=n(-a+a)=n \equiv 0 \mod{n}$ and $a+(n-a)=n+(a-a)=n \equiv
      0 \mod{n}$. So $(\faktor{\Z}{n\Z}, +_n)$ is a group, with identity
      element $0 \mod{n}$ and inverse element $-a \mod{n}$ for each $a
      \in \faktor{\Z}{n\Z}$.
  \end{enumerate}
\end{example}

\begin{definition}
  We call a group $G$ under a binary operation  $\ast$  \textbf{Abelian}, or
  \textbf{commutative} if for every $a,b \in G$,  $ab=ba$.
\end{definition}

\begin{example}\label{example_1.3}
  \begin{enumerate}
    \item[(1)] The above examples of groups are also examples of abelian groups.

    \item[(2)] Consider $\com{(\faktor{\Z}{n\Z})}{\{0\}}$ under multiplication
      modulo $n$, $\cdot_n$  (abbreviated as $\cdot$). This is not a group as
      not every element has an inverse. Specifically, take $n=6$, then in
      $\faktor{\Z}{6\Z}$, $2 \cdot 3=6 \equiv 0 \mod{6} \notin
      \com{(\faktor{\Z}{n\Z})}{\{0\}}$. However, one can still impose a group
      structure with modular multiplication.

      Define the set $\Uc{(\faktor{\Z}{n\Z})}=\{a \in \faktor{\Z}{n\Z} : (a,n)=1\}$,
      that is it is the set of all integers modulo $n$, coprime with $n$. We have
      that $\Uc{(\faktor{\Z}{n\Z})}$ is closed under $\cdot$. Notice that if
      $(a,n)=1$, and $(b,n)=1$, then $(ab,n)=1$. Also notice that
      $\Uc{(\faktor{\Z}{n\Z})} \subseteq \faktor{\Z}{n\Z}$, and so inherits
      associativity. Moreover, notice that $(1,n)=1$, and $a1=1a=a \in
      \Uc{(\faktor{\Z}{n\Z})}$, so $1 \mod{n}$ is the identity of the set. Now for
      any $a \in \Uc{(\faktor{\Z}{n\Z})}$, since $(a,n)=1$, there exist $b,m \in
      \Z$ such that $ab+mn=1$, that is $ab \equiv 1 \mod{n}$, and moreover notice
      that if $(ab,n)=1$ and $(a,n)=1$, then $(b,n)=1$, thus $b \in
      \Uc{(\faktor{\Z}{n\Z})}$, and is an inverse of $a$. Thus we have shown that
      $\Uc{(\faktor{\Z}{n\Z})}$ is a group under $\cdot$. We call the group the
      \textbf{group of units, modulo $n$}, or simply the \textbf{unit group $\mod{n}$}.
      Moreover, we see that this group is commutative. We observe that it
      $n=p$ is a prime, then  $\Uc{(\faktor{\Z}{pZ})}=\faktor{\Z}{p\Z}$. The concept
      of a ``group of units'' can be expanded to more general algebraic structures
      than $\faktor{\Z}{n\Z}$. Consider $\com{(\faktor{\Z}{n\Z})}{\{0\}}$ under multiplication
      modulo $n$, $\cdot_n$  (abbreviated as $\cdot$). This is not a group as
      not every element has an inverse. Specifically, take $n=6$, then in
      $\faktor{\Z}{6\Z}$, $2 \cdot 3=6 \equiv 0 \mod{6} \notin
      \com{(\faktor{\Z}{n\Z})}{\{0\}}$. However, one can still impose a group
      structure with modular multiplication.

      Define the set $\Uc{(\faktor{\Z}{n\Z})}=\{a \in \faktor{\Z}{n\Z} : (a,n)=1\}$,
      that is it is the set of all integers modulo $n$, coprime with $n$. We have
      that $\Uc{(\faktor{\Z}{n\Z})}$ is closed under $\cdot$. Notice that if
      $(a,n)=1$, and $(b,n)=1$, then $(ab,n)=1$. Also notice that
      $\Uc{(\faktor{\Z}{n\Z})} \subseteq \faktor{\Z}{n\Z}$, and so inherits
      associativity. Moreover, notice that $(1,n)=1$, and $a1=1a=a \in
      \Uc{(\faktor{\Z}{n\Z})}$, so $1 \mod{n}$ is the identity of the set. Now for
      any $a \in \Uc{(\faktor{\Z}{n\Z})}$, since $(a,n)=1$, there exist $b,m \in
      \Z$ such that $ab+mn=1$, that is $ab \equiv 1 \mod{n}$, and moreover notice
      that if $(ab,n)=1$ and $(a,n)=1$, then $(b,n)=1$, thus $b \in
      \Uc{(\faktor{\Z}{n\Z})}$, and is an inverse of $a$. Thus we have shown that
      $\Uc{(\faktor{\Z}{n\Z})}$ is a group under $\cdot$. We call the group the
      \textbf{group of units, modulo $n$}, or simply the \textbf{unit group $\mod{n}$}.
      Moreover, we see that this group is commutative. We observe that it
      $n=p$ is a prime, then  $\Uc{(\faktor{\Z}{pZ})}=\faktor{\Z}{p\Z}$. The concept
      of a ``group of units'' can be expanded to more general algebraic structures
      than $\faktor{\Z}{n\Z}$.

    \item[(3)] The vector space axioms for some vector space $V$ specify
      that under vector addition  $+$,  $(V,+)$ forms a group.

    \item[(4)] Let $(A, \ast)$, $(B,\cdot)$ be groups under binary
      operations $\ast$ and  $\cdot$. Consider the product $A \times B$
      and take the map $\circ:(a_1,b_1)(a_2,b_2) \rightarrow (a_1 \ast a_2, b_1
      \cdot b_2)$. Then $\circ$ is a binary operation on  $A \times B$.
      Then  $(A \times B, \circ)$ forms a group. We have that since $A$
      and  $B$ is closed, then so is  $A \times B$. Furthermore, by
      associativity of  $\ast$ and  $\cdot$,  $((a_1,b_1) \circ (a_2,b_2))
      \circ (a_3,b_3)=(a_1,b_1) \circ ((a_2,b_2) \circ (a_3,b_3))$; making
      $\circ$ associative. Now if  $e_1$ and $e_2$ are the identities of
      $A$ and  $B$ respevively, then  $(e_1,e_2)$ is the identiy for $A
      \times B$; finding the inverse of an element  $(a,b)$ follow
      similarly.
  \end{enumerate}
\end{example}

\begin{theorem}\label{theorem_1.2.1}
  Let $G$ be a group under a binary operation  $\ast$ then the identy of  $G$
  is unique, and the inverse of  $a \in G$ is unique.
\end{theorem}
\begin{proof}
  Suppose there exists $e,f \in G$ such that for any  $a \in G$  $ae=ea=a$ and
  $af=fa=a$. Then we have that $fe=e$ and  $ef=fe=f$; thus  $e=f$.

  Now let  $a \in G$, and suppose  $a$ has inverses  $b,c \in G$, then
  $ab=ba=e$ and  $ac=ca=e$, where  $e$ is the identity of  $G$. Then we have
  $b=be=b(ac)=(ba)c=ec=c$, thus $b=c$.
\end{proof}
\begin{remark}
  Since the inverse of an element $a$ is unique, we will now denote it
  $\inv{a}$.
\end{remark}
\begin{corollary}
  $\inv{(\inv{a})}=a$ and $\inv{(ab)}=\inv{b}\inv{a}$
\end{corollary}
\begin{proof}
  Since inverses are unique, the $\inv{a} \in G$ has the unique inverse
  $\inv{(\inv{a})}$. Then  taking $a\inv{a}=e$, applying inverses, we get
  $a(\inv{a}\inv{(\inv{a})})=e\inv{(\inv{a})}$, $a=\inv{(\inv{a})}$.

  Now Let $a,b \in G$, then $ab\inv{(ab)}=e$. Applying the inverse of $a$ on
  the right to both sides, we get  $b\inv{(ab)}=\inv{a}$; agian with the
  inverse of $b$ yields  $\inv{(ab)}=\inv{b}\inv{a}$.
\end{proof}

\begin{theorem}[Generalized Associativity]\label{theorem_1.2.2}
  Let $G$ be a group under a binary operation  $\ast$, then for any  $a_1,
  a_2, \dots, a_n \in G$, the product $a_1 \ast a_1 \ast \dots \ast a_n$ is
  independent of the ordering of any brackets.
\end{theorem}
\begin{proof}
  By induction on $n$, for $n=3$, we have the elements $a_1,a_2,a_3
  \in G$, and by the group law of associativity, we are gauranteed
  that $a_1 \ast (a_2 \ast a_3)=(a_1 \ast a_2) \ast a_3$.

  Now, for $n \geq 3$, suppose that we can bracket the elements
  independent of order, and consider the expression
  \begin{equation*}
    a_1 \ast (a_2 \ast (a_3 \ast ( \dots \ast (a_n \ast a_{n+1}))))
  \end{equation*}
  we may rebracket this expression as:
  \begin{equation*}
    (a_1 \ast (a_2 \ast (a_3 \ast ( \dots \ast a_n)))) \ast a_{n+1}
  \end{equation*}
  which by hypothesis, can be rebracketed as
  \begin{equation*}
    ((a_1 \ast \dots \ast a_k) \ast (a_{k+1} \ast \dots \ast a_n)) \ast a_{n+1}
  \end{equation*}
  which by the associative law, just gives us the result
  \begin{equation*}
    a_1 \ast \dots \ast a_{n+1}=a_1 \ast (a_2 \ast (a_3 \ast ( \dots \ast (a_n \ast a_{n+1}))))
  \end{equation*}
  which completes the proof
\end{proof}

\begin{theorem}[The Cancellation Laws]\label{theorem_1.2.3}
  Let $G$ be a group under a binary operation  $\ast$. Then for  $a,b,c \in
  G$, we have:
  \begin{enumerate}
    \item[(1)] $ab=ac$ implies  $b=c$.

    \item[(2)] $ba=ca$ implies  $b=c$.
  \end{enumerate}
\end{theorem}
\begin{proof}
  Suppose that $ab=ac$, then applying the inverse of $a$ on the left, we get
  $(\inv{a}a)b=(\inv{a}a)c$, hence $eb=ec$, thus  $b=c$. Similarly, we get
  $b=c$ if we apply  $\inv{a}$ to the right in the equation $ba=ca$.
\end{proof}
\begin{corollary}
  For $x,y \in G$, the equations  $ax=b$ and  $ya=b$ have unique solutions.
\end{corollary}
\begin{proof}
  We have $x=b\inv{a}$ and $y=\inv{a}b$. Since inverses are unique, so are the
  solutions $x$ and $y$.
\end{proof}

\begin{definition}
  Let $G$ be a group under a binary opeartion  $\ast$. For any $a \in G$, and
  $n \in \Z^+$, we define the  \textbf{$n$-th power} of $a$ to be:
  \begin{equation*}
    a^n=\underbrace{a \ast \dots \ast a}_{n \text{ times.}}
  \end{equation*}
  We define $a^0=e$ and  $a^{-n}=(\inv{a})^n$.
\end{definition}

\begin{proposition}\label{proposition_1.2.4}
  Let $G$ be a group under a binary operation  $\ast$, and let  $a \in G$ and
  $m,n \in \Z^+$. Then:
  \begin{enumerate}
    \item[(1)] $a^ma^n=a^{m+n}$.

    \item[(2)] $(a^m)^n=a^{mn}$.
  \end{enumerate}
\end{proposition}
\begin{proof}
  We have by definition that $a^ma^n=\underbrace{a \ast \dots \ast a}_{m
  \text{ times.}}\underbrace{a \ast \dots \ast a}_{n \text{ times.}}=
  \underbrace{a \ast \dots \ast a}_{m+n \text{ times.}}=a^{m+n}$.

  Likewise, $(a^m)^n=\underbrace{a^m \ast \dots \ast a^m}_{n \text{ times.}}=
  \underbrace{a \ast \dots \ast a}_{nm \text{ times.}}=a^{mn}$.
\end{proof}

We can now, unless context isn't clear enough, drop all mention to the binary
operation of a group.

\begin{definition}
  We definet the \textbf{order} of a group $G$ to be the number of
  elements of  $G$  (i.e. the cardinality of $G$) and denote it $|G|$.
  If $G$ is infinite, we say that $G$ has \textbf{infinite order} and
  write $|G|=\infty$; otherwise, we say that $G$ has \textbf{finite
  order}.
\end{definition}

\begin{definition}
  Let $G$ be a group and $a \in G$. We define the \textbf{order} of
  $a$ to be the smallest postivie such integer $n \in \Z^+$ for which
  $a^n=e$, and write $|a|=n$. If there is no such positive integer,
  then we say that $a$ has \textbf{infinite order} and write
  $|a|=\infty$; otherwise we say that $a$ is of \textbf{finite order}.
\end{definition}

We conclude the section with more examples and one last definition.

\begin{proposition}\label{proposition_1.2.5}
  Let $G$ be a group, suppose that  $a,b \in G$ are elements with
  order $|a|=m$ and  $|b|=n$, for  $m,n \in \Z^+$. Suppose further
  that $ab=ba$. Then $|ab|=[m,n]$, the least common multiple of $m$
  and  $n$.
\end{proposition}
\begin{proof}
  Let $|ab|=k$ for $k \in \ \Z^+$. Since  $a$ and  $b$ commute, we get
  $a^k=e$ and  $b^k=e$. Now, by the division theorem, there are integers
  $q_1,q_2,r_1,r_2 \in \Z^+$ such that $k=q_1m+r_1$ and $k=q_2n+r_2$. Then we
  get that $a^k=a^{q_1m+r_1}=a^{q_1m}a^{r_1}=a^{r_1}=e$. Since $|a|=m$,
  this makes $r_1=0$. Similarly, we get that $b^k=b^{r_2}=e$ implies $r_2=0$.
  Therefore, $k=q_1m=q_2n$; moreover, $|ab|=k$ is minimal by definition,
  thus we get $|ab|=k=[m,n]$.
\end{proof}
\begin{corollary}
  For $a_1, \dots, a_k \in G$, with $|a_i|=n_i$ and $a_ia_j=a_ja_i$ for
  all  $1 \leq i,j \leq n$, then  $|a_1a_2 \dots a_k|=[n_1, \dots, n_k]$.
\end{corollary}
\begin{proof}
  By induction, when $k=1$, the case is trivial. Now, for  $k=2$, the result
  follows by the above theorem. Now suppose that $|a_1 \dots a_k|=[n_1,
  \dots, n_k]$. Take $a_{k+1} \in G$ with $|a_{k+1}|=n_{k+1}$ and $(a_1
  \dots a_k)a_{k+1}=a_{k+1}(a_1 \dots a_k)=a_1 \dots a_ka_{k+1}$. Then, by the
  above theorem, we get $|a_1 \dots a_ka_{k+1}|=[[n_1, \dots, n_k],
  n_{k+1}]=[n_1, \dots, n_k, n_{k+1}]$.
\end{proof}

\begin{example}
  \begin{enumerate}
    \item[(1)] $|\langle e \rangle|=1$
    \item[(1)] In any group $G$,  $|e|=1$, and if $a \in G$ has
      $|a|=1$, then necessarily, $a=e$. That is in any group, the only
      element of order  $1$ is the identity.

    \item[(2)] The additive groups $\Z$,  $\Q$,  $\R$, and  $\C$ have
      infinite order, and their nonzero elements also have infinite order.

    \item[(3)] The multiplicative group $\C^*$ has infinite order, morover,
      since $i^4=1$ and $i^3=-i$, $|i|=4$ (not to be confused with the
      modulus of the complex number $i$).

    \item[(4)] $|\faktor{\Z}{n\Z}|=n$, and every element is of finite
      order. $|\Uc{(\faktor{\Z}{n\Z})}|=\phi(n)$, where $\phi$ is the Euler
      totient function. Every element of $\Uc{(\faktor{\Z}{n\Z})}$ also has
      finite order.
  \end{enumerate}
\end{example}

\begin{definition}
  Let $G=\{g_1, \dots, g_n\}$ be a finite group of order $n$ with $g_1=e$. We
  define the \textbf{Cayley table}, or \textbf{multiplication table} of $G$ to
  be the  $n \times n$ matrix defined by the entries $(g_{ij})$ where
  \begin{equation*}
    g_{ij}=g_ig_j
  \end{equation*}
  for $1 \leq i,j \leq n$.
\end{definition}
