\section{Outer and Inner Approximations}

\begin{theorem}[The Outer Approximation Theorem]\label{8.4.1}
    LEt $E \subseteq \R$. Then the following are true.
    \begin{enumerate}
        \item[(1)] $E$ is measurable.

        \item[(2)] For every $\e>0$, there is an open set  $U$, containing  $E$
            for which $m^\ast(\com{U}{E})<\e$

        \item[(3)] There exists a $G_\d$ set  $G$, containing $E$ for which
            $m^\ast(\com{G}{E})=0$.
    \end{enumerate}
\end{theorem}
\begin{proof}
    Suppose that $E$ is measurable, and let  $\e>0$. IF  $m^\ast(E)$ is finite,
    then by definition of outer measure, there exists a countable collection of
    open bounded intervals $\{I_k\}$ covering $E$, for which
    \begin{equation*}
        \sum{l(I_k)}<m^\ast(E)+\e
    \end{equation*}
    Define now, $U=\bigcup{I_k}$, then $U$ is open in  $\R$, and  $E \subseteq
    \R$. Again, by the definition of outer measure, we have
    \begin{equation*}
        m^\ast(U) \leq \sum{l(I_k)}<m^\ast(E)+\e
    \end{equation*}
    so that
    \begin{equation*}
        m^\ast(U)-m^\ast(E)<\e
    \end{equation*}
    Now, since $E$ is measurable, and  $E \subseteq U$, by the excision property
    we get $m^\ast(\com{U}{E})<\e$.

    Now, if $m^\ast(E)$ is infinite, then there exists a disjoint countable
    collection $\{E_k\}$ of measurable sets, each of finite outer measure, for
    which $E=\bigcup{E_k}$. Then for any $k \in \Z^+$, there exists an open set
     $U_k$ with  $E_k \subseteq U_k$ for which
     \begin{equation*}
         m^\ast(\com{U_k}{E_k})<\frac{\e}{2^k}
     \end{equation*}
     Now, let $U=\bigcup{U_k}$, so that $U$ is open and  $E \subseteq U$.
     Moreover, we get $\com{U}{E} \subseteq \bigcup{\com{U_k}{E_k}}$; therefore,
     by monotonicity, we get
     \begin{equation*}
         m^\ast(\com{U}{E}) \leq
         \sum{m^\ast(\com{U_k}{E_k})}<\sum{\frac{\e}{2^k}}=\e
     \end{equation*}

     Now, suppose that the second assertion holds for $E$. Then for any  $k \in
     \Z^+$, choose an open set  $U_k$ containing  $E$ for which
     \begin{equation*}
         m^\ast(\com{U_k}{E})<\frac{1}{k}
     \end{equation*}
     Define $G=\bigcap{U_k}$. Then $G$ is a $G_\d$ set and $E \subseteq G$.
     Moreover, for each $k$, we have $\com{G}{E} \subseteq \com{U_k}{E}$ so by
     monotonicity
     \begin{equation*}
         m^\ast(\com{G}{E}) \leq m^\ast(\com{U_k}{E})<\frac{1}{k}
     \end{equation*}
     Then as $k \xrightarrow{} \infty$, $m^\ast(\com{G}{E}) \xrightarrow{} 0$.

     Finally, suppose that the third assertion holds for $E$. Now, since sets of
     measure  $0$ are measurable, and  $G_\d$ are measurable, we get
     \begin{equation*}
         E=G \cap \com{\R}{(\com{G}{E})}
     \end{equation*}
     is measurable, by theorem \ref{8.3.6}.
\end{proof}

\begin{theorem}[The Inner Approximation Theorem]\label{8.4.1}
    LEt $E \subseteq \R$. Then the following are true.
    \begin{enumerate}
        \item[(1)] $E$ is measurable.

        \item[(2)] For every $\e>0$, there is an closed set $F$, contained in $E$
            for which $m^\ast(\com{E}{F})<\e$

        \item[(3)] There exists a $F_\s$ set  $F$, contained in $E$ for which
            $m^\ast(\com{E}{F})=0$.
    \end{enumerate}
\end{theorem}
\begin{proof}
    This follows by DeMorgan's laws on the outer approximation theorem.
\end{proof}

\begin{theorem}\label{8.4.3}
    Let $E$ be measurable of finite outer measure. Then for every  $\e>0$, there
    exists a finite disjoint collection of open intervals  $\{I_k\}_{k=1}^n$ for
    which if $U=\bigcup_{k=1}^n{I_k}$, then
    \begin{equation*}
        m^\ast(\com{E}{U})+m^\ast(\com{U}{E})<\e
    \end{equation*}
\end{theorem}
\begin{proof}
    Since $E$ is measurable, there exists an open set  $V$, containing  $E$ and
    for which
    \begin{equation*}
        m^\ast(\com{V}{E})<\frac{\e}{2}
    \end{equation*}
    Then by the excision property, we observe that since $E$ has finite outer
    measure, then $V$ must also have finite outer measure. Now, let
    $V=\bigcup{I_k}$ where $\{I_k\}$ is a countable disjoint collection of open
    intervals. Therefore, by lemma \ref{8.3.3}, and monotonicity we get
    \begin{equation*}
        \sum_{k=1}^n{l(I_k)}=m^\ast\Big{(} \bigcup_{k=1}^n{I_k} \Big{)} \leq
        m^\ast(V)
    \end{equation*}
    Then as $n \xrightarrow{} \infty$, we see the sum $\sum{l(I_k)}$ is finite.
    Then choosing $n \in \Z^+$ for which
    \begin{equation*}
        \sum_{k=n+1}^{\infty}{l(I_k)}<\frac{\e}{2}
    \end{equation*}
    and defining
    \begin{equation*}
        U=\bigcup_{k=1}^n{I_k}
    \end{equation*}
    we get $\com{U}{E} \subseteq \com{V}{E}$ so by monotonicity,
    \begin{equation*}
        m^\ast(\com{U}{E}) \leq m^\ast(\com{V}{E})<\frac{\e}{2}
    \end{equation*}
    on the other hand, since $E \subseteq V$, we get  $\com{E}{U} \subseteq
    \com{V}{U}=\bigcup_{k=n+1}^\infty{I_k}$. Therefore by definition of outer
    measure
    \begin{equation*}
        m^\ast(\com{E}{U}) \leq \sum_{k=n+1}^\infty{l(I_k)}<\frac{\e}{2}
    \end{equation*}
    This gives us
    \begin{equation*}
        m^\ast(\com{U}{E})+m^\ast(\com{E}{U})<\frac{\e}{2}+\frac{\e}{2}=\e
    \end{equation*}
\end{proof}
