\section{Tensor Products of Modules}

\begin{definition}
    Let $S$ be a ring, and $R$ a subring of $S$. Let $f:S
    \xrightarrow{} R$ a ring homomorphism with $f(1_S)=1_R$ and let $N$ be an
    $R$-module, with the scalar multiplication defined by $rn=f(r)n$. We call
    $N$ a \textbf{restriction of scalars} from $S$ to $R$. We call $S$ an
    \textbf{extension} of $R$.
\end{definition}

\begin{example}\label{example_4.8}
    \begin{enumerate}
        \item[(1)] Let $S$ a  ring and $R$ a subring of $S$. We can
            restrict an $S$-module $N$ to be an $R$-module, but the reverse is
            not in general true. Consider $\Z$ as a $\Z$-module, then $\Z$ cannot
            be a $\Q$-module, for if it is, then the element $\frac{1}{2}1=z$ is
            an element of $\Z$ for which $z+z=1$, which cannot happen. Notice
            however that $\Z$ can be imbedded into a $\Q$-module, namely $\Q$ itself.

        \item[(2)] Consider al possible $\Z$-module homomorphisms of
            $\faktor{\Z}{2\Z}$ into a $\Q$-module $V$. Since $\Q$ is a field $V$
            is a vector space over $\Q$, and every element has infinite order.
            Since the elements of $\faktor{\Z}{2\Z}$ are of finite order, then
            $x \xrightarrow{} 0$ for all $x \in \faktor{\Z}{2\Z}$. This shows
            that $\faktor{\Z}{2\Z}$ cannot be imbedded into a $\Q$-module.
    \end{enumerate}
\end{example}

\begin{definition}
    Let $R$ and $S$ be rings, and let $N$ be an $R$-module. Consider the map $S
    \times N \xrightarrow{} N$ defined by $(s,n) \xrightarrow{} sn$. Consider
    $H$ the subgroup generated by all elements of the form
    $(s_1+s_1,n)-(s_1,n)-(s_2,n)$, $(s,n_1+n_2)-(s,n_1)-(s,n_2)$, and
    $(sr,n)-(s,rn)$ for all $s,s_1,s_2 \in S$, $r \in R$, and $n,n_1,n_2 \in N$.
    We define the \textbf{tensor product} of $S$ together with  $N$ over $R$ to
    be the quotient group $\faktor{N}{H}$. We denote this tensor product by $S
    \otimes_R N$. We denote the elements of $S \otimes_R N$ as $s \otimes n$ and
    call them \textbf{simple tensors}. We call sums of simple tensors
    \textbf{tensors}.
\end{definition}

\begin{lemma}\label{4.4.1}
    Let $R$ and $S$ be rings, and let $N$ be an $R$-module. Then the following
    relations hold on the tensor product $S \otimes_R N$
    \begin{enumerate}
        \item[(1)] $(s_1+s_2) \otimes n=s_1 \otimes n+s_2 \otimes n$.

        \item[(2)] $s \otimes (n_1+n_2)=s \otimes n_1+s \otimes n_2$.

        \item[(3)] $sr \otimes n=s \otimes rn$.
    \end{enumerate}
\end{lemma}
\begin{proof}
    Since $S \otimes_R N$ is the quotient group of $N$ by $H$, where $H$ the
    subgroup generated by all elements of the form $(s_1+s_1,n)-(s_1,n)-(s_2,n)$,
    $(s,n_1+n_2)-(s,n_1)-(s,n_2)$, and $(sr,n)-(s,rn)$; that is: $H=\langle
    (s_1+s_1,n)-(s_1,n)-(s_2,n), (s,n_1+n_2)-(s,n_1)-(s,n_2), (sr,n)-(s,rn)
    \rangle$. Then the tensors $s \otimes n$ are just cosets of $H$ in $N$, and
    the relations defining $H$ give us the relations on tensors.
\end{proof}

\begin{theorem}\label{4.4.2}
    Let $S$ be a ring, and $R$ a subring of $S$. Let $N$ be an
    $R$-module, then the tensor product $S \otimes_R N$ is in $S$-module under
    the action
    \begin{equation*}
        s\sum{s_i \otimes n_i}=\sum{(ss_i) \otimes n_i}
    \end{equation*}
    where the sums are finite.
\end{theorem}
\begin{proof}
    Let $s' \in S$, then
    \begin{align*}
        (s'(s_1+s_2),n)-(s's_1,n)-(s's_2,n) &= (s's_1+s's_2,n)-(s's_1,n)-(s's_2,n) \\
        (s's,n_1n_2)-(s's,n_1)-(s's,n_2)    \\
        (s'(sr),n)-(s's,rn) &=  ((s's)r,n)-(s's,rn) \\
    \end{align*}
    each belong to the generating set
    of $H=\langle (s_1+s_1,n)-(s_1,n)-(s_2,n), (s,n_1+n_2)-(s,n_1)-(s,n_2),
    (sr,n)-(s,rn) \rangle$. Since elements of $H$ are sums of elements in $H$,
    for every $\sum{(s_i,n_i)} \in H$, we get $\sum{(s's_i,n_i)} \in H$

    Now, suppose that $\sum{s_i \otimes n_i}=\sum{s'_i \otimes n'_i}$ are
    representatives of the same tensor in $S \otimes_R N$. Then
    $\sum{(s_i,n_i)}-\sum{(s'_i,n'_i)} \in H$ so that $\sum{ss_i \otimes
    n}=\sum{ss'_i \otimes n} \in S \otimes_R N$. This makes the action well
    defined.

    Finally, observe that for every tensor $s_i \otimes n_i \in S \otimes_R N$,
    that the following relation holds, by lemma \ref{4.4.1}
    \begin{equation*}
        (s+s')(s_i \otimes n_i)=s(s_i \otimes n_i)+s'(s_i \otimes n_i)
    \end{equation*}
    The other relations for modules also hold by lemma \ref{4.4.1}.
\end{proof}

\begin{definition}
    Let $S$ be a rings and $R$ a subring of $S$, and let $N$ be an
    $R$-module. We call $S \otimes_R N$, as a module, the \textbf{left
    $S$-module obtained by an extension of scalars} from $N$.
\end{definition}

\begin{theorem}\label{4.4.3}
    Let $R$ a subring of a ring $S$, and let $N$ be a left $R$-module. Let $\i:N
    \xrightarrow{} S \otimes_R N$ be the $R$-module homomorphism defined by $\i:
    \xrightarrow{} 1 \otimes n$. If $L$ is any left $S$-module, and $\phi:N
    \xrightarrow{} L$ is an $R$-module homomorphism, then there exists a unique
    $S$-module homomorphism $\Phi:S \otimes_R N \xrightarrow{} L$ for which
    $\phi=\Phi \circ \i$. That is, the following diagram commutes.
    \[\begin{tikzcd}
        N & {S \otimes_R N} \\
        & L
        \arrow["\iota", from=1-1, to=1-2]
        \arrow["\Phi", from=1-2, to=2-2]
        \arrow["\phi"', from=1-1, to=2-2]
    \end{tikzcd}\]
    Conversely, if $\Phi:S \otimes_R N \xrightarrow{} L$ is an $S$-module
    homomorphism, then the map $\phi=\Phi \circ \i$ is an $R$-module
    homomorphism.
\end{theorem}
\begin{proof}
    Let $\phi:N \xrightarrow{} L$ an $R$-module homomorphism. By the uiversal
    property of free modules, there exists a $\Z$-module homomorphism from the
    free $\Z$-module $F(S \times N)$ to  $L$ defined by $(s,n) \xrightarrow{}
    s\phi(n)$. Since $\phi$ is an $R$-module homomorphism, the generators of $S
    \otimes_R N$, as a quotient group, map to $0$ in $L$. Therefore, there
    exists a well define $\Z$-module homomorphism $\Phi:S \otimes_R N
    \xrightarrow{} L$ taking $s \otimes n \xrightarrow{} s\phi(n)$. Then for
    simple tensors, we get $s'\Phi(s \otimes
    n)=s'(s\phi(n))=(s's)\phi(n)=\Phi(s'(s \otimes n))$ for all $s' \in S$.
    Since $\Phi$ is additive, we get an $S$-module homomorphism.

    Now, $S \otimes_R N$, as an $S$-module, is generated by all $1 \otimes n$,
    and any $S$-module homomorphism is uniquely determined by its value on these
    elements. Now, observe that $\Phi(1 \otimes n)=1\phi(n)=\phi(n)$, so
    $\phi=\Phi \circ \i$ and $\Phi$ is uniquely determined by $\phi$.

    Now, if we have the map $\phi=\Phi \circ \i$, where $\Phi$ is defined as
    above, and $\i:n \xrightarrow{} 1 \otimes n$, then since $\Phi$ is an
    $S$-module homomorphsm, and $\i$ is an $R$-module homomorphism, then $\phi$
    must be an $R$-module homomorphism.
\end{proof}
\begin{corollary}
    $\faktor{N}{\ker{\i}}$ is the largest unique quotient of $N$ that can be
    imbedded into any $S$-module. In particular, $N$ can be imbedded as an
    $R$-module into some left $S$-module if, and only if $\i$ is 1--1.
\end{corollary}
\begin{proof}
    We have that $\faktor{N}{\ker{\i}}$ is mapped into $S \otimes_R N$
    injectively. Now, let $\phi$ be an 1--1 $R$-module taking
    $\faktor{N}{\ker{\i}} \xrightarrow{} L$. Then $\ker{\i} \xrightarrow{} (0)$
    by $\phi$ so that $\ker{\i} \leq \ker{\phi}$. This makes
    $\faktor{N}{\ker{\phi}}$ a quotient of $\faktor{N}{\ker{\i}}$.
\end{proof}

\begin{example}\label{example_4.9}
    \begin{enumerate}
        \item[(1)] Let $N$ be an $R$-module, and take $\phi:N \xrightarrow{} N,
            n \xrightarrow{} n$. Then $R \otimes_R N \simeq R$ via this map. In
            particular, if $A$ is any Abelian group, then $\Z \otimes_\Z A
            \simeq A$.

        \item[(2)] Let $A$ be a finite Abelian group of order $n$. THen $\Q
            \otimes_\Z A=(0)$. Observe that $1 \otimes 0=1 \otimes (0+0)=1
            \otimes 0+1 \otimes 0$, so that $1 \otimes 0=0$. Then for any $q
            \otimes a \in \Q \otimes_\Z A$, we have $q \otimes a=(\frac{q}{n}n)
            \otimes a=\frac{q}{n} \otimes na=\frac{q}{n} \otimes 0=\frac{q}{n}(1
            \otimes 0)=\frac{q}{n}0=0$; since $na=0$. Moreover, the map $\i:A
            \xrightarrow{} \Q \otimes_\Z A$ is the zero map, so that if $A \neq
            \langle 0 \rangle$, any map $\phi:\Q \xrightarrow{} A$ is also the
            zeromap.

        \item[(3)] Let $R$ be a ring, and consider the free module $R^n$, of
            rank $n$. Then $S \otimes_R R^n \simeq S^n$, the free module over
            $S$ of rank $n$. Then $\Q \otimes_\Z \Z^n \simeq \Q^n$. Moreover,
            observe that $\Z^n$ is a subgroup of $\Q^n$.

        \item[(4)] Let $K$ be a field, and $F$ a subfield of $K$. Let $V$ an
            $n$-dimensional vector space over $F$. Then $K \otimes_F V \simeq
            K^n$. We recall that $V \simeq F^n$.

        \item[(5)] Let $R$ be a commutative ring with identity, and $G$ a finite
            group. Let $H \leq G$. For any $RH$-module $N$, we define the
            \textbf{induced module} $RG \otimes_{RH} N$ obtained as an extension
            of scalars from $N$ as an $RG$-module.
    \end{enumerate}
\end{example}

\begin{definition}
    Let $R$ be a ring, and $N$ a left $R$-module, and $M$ a right $R$-module.
    THe quotient of the free $\Z$-module on $M \times N$ by the subgroup
    \begin{equation*}
        H=\langle (m_1+m_1,n)-(m_1,n)-(m_2,n), (m,n_1+n_2)-(m,n_1)-(m,n_2), (mr,n)-(m,rn) \rangle
    \end{equation*}
    is called the \textbf{tensor product} of $M \times N$ over $R$, and denoted
    $M \otimes_R N$. We call the elements $m \times n$ \textbf{simple tensors},
    and we call sums of simple tensors \textbf{tensors}.
\end{definition}

\begin{lemma}\label{4.4.4}
    Let $R$ be a ring, and $N$ and $M$ left and right $R$-modules, respectively.
    Then the following relations hold.
    \begin{enumerate}
        \item[(1)] $(m_1+m_2) \otimes n=m_1 \otimes n+m_2 \otimes n$.

        \item[(2)] $m \otimes (n_1+n_2)=m \otimes n_1+m \otimes n_2$.

        \item[(3)] $mr \otimes n=m \otimes rn$.
    \end{enumerate}
\end{lemma}
\begin{proof}
    The proof is identical to that of lemma \ref{4.4.1}.
\end{proof}

\begin{definition}
    Let $R$ be a ring, and let $M$ and $N$ be right and left $R$-modules,
    respectively, and let $L$ be an Abelian group. We call the map $\phi:M
    \times N \xrightarrow{} L$ \textbf{$R$-balanced} or \textbf{middle linear}
    with resepct to $R$ if for all $m_1,m_2 \in M$, and $n_1,n_2 \in N$, and $r
    \in R$
    \begin{enumerate}
        \item[(1)] $\phi(m_1+m_2,n)=\phi(m_1,n)+\phi(m_2,n)$.

        \item[(2)] $\phi(m,n+1+n_2)=\phi(m,n_1)+\phi(m,n_2)$

        \item[(3)] $\phi(mr,n)=\phi(m,rn)$.
    \end{enumerate}
\end{definition}

\begin{lemma}\label{4.4.5}
    Let $R$ be a ring, and $M$ and $N$ right, and left $R$-modules respectively.
    The map $\i:M \times N \xrightarrow{} M \otimes_R N$ defined by $\i:(m,n)
    \xrightarrow{} m \otimes n$ is $R$-balanced.
\end{lemma}
\begin{proof}
    This is by definition of the tensor product.
\end{proof}

\begin{theorem}\label{4.4.6}
    Let $R$ be a ring with identity, and $M$ and $N$ right and left modules,
    respectively. Let $\i:M \times N \xrightarrow{} M \otimes_R N$ the
    $R$-balanced map defined by $\i:(m,n) \xrightarrow{} m \otimes n$. Then the
    following are true.
    \begin{enumerate}
        \item[(1)] If $\Phi:M \otimes_R N \xrightarrow{} L$ is a group
            homomorphsm and $L$ an Abelian group, then the map $\phi=\Phi \circ
            \i$ is $R$-balanced from $M \times N \xrightarrow{} L$.

        \item[(2)] If $L$ is an Abelian group, and $\phi:M \times N
            \xrightarrow{} L$ is an $R$-balanced map, then there exists a unique
            group homomorphism $\Phi:M \otimes_R N \xrightarrow{} L$ for wich
            $\phi=\Phi \circ \i$.
    \end{enumerate}
\end{theorem}
\begin{proof}
    We first have that if $\phi(m,n)=\Phi(m \otimes n)$, then
    $\phi(m_1+m_2,n)=\Phi((m_1+m_2) \otimes n)=\Phi(m_1 \otimes n)+\Phi(m_2
    \otimes n)=\phi(m_1,n)+\phi(m_2,n)$. Similarly, the rest of the conditions
    for $R$-balanced maps hold for $\phi$, making it $R$ balanced.

    Conversely by the universal propery, $\phi$ defines a unique $\Z$-module
    homomorphism $\tilde{\phi}:F(M \times N) \xrightarrow{} L$ for wich
    $\tilde{\phi}(m,n)=\phi(m,n) \in L$. Now, if $\phi$ is $R$-balanced, then
    $\tilde{\phi}$ takes $(m,n) \xrightarrow{} 0$, that is, we get
    $\tilde{\phi}((mr,n)-(m,rn))=0$. Then the subgroup $H$ defined by
    \begin{equation*}
        H=\langle (m_1+m_1,n)-(m_1,n)-(m_2,n), (m,n_1+n_2)-(m,n_1)-(m,n_2), (mr,n)-(m,rn) \rangle
    \end{equation*}
    is a subgroup of $\tilde{\phi}(F(M \times N))$. Then $\tilde{\phi}$ indues a
    homomorphism $\Phi:M \otimes_R N \xrightarrow{} L$. By definition we have
    $\Phi(m \otimes n)=\tilde{\phi}(m,n)=\phi(m,n)$.
\end{proof}
\begin{corollary}
    Let $D$ be an Abelian group, and $\i':M \times N \xrightarrow{} D$ such that
    \begin{enumerate}
        \item[(1)] $\i'(M \times N)$ generates $D$ as a group.

        \item[(2)] Every $R$-balanced map on $M \times N$ factors through $\i'$.
    \end{enumerate}
    Then there exists an isomorphism $f:M \otimes_R N \xrightarrow{} D$ with
    $\i'=f \circ \i$.
\end{corollary}
\begin{proof}
    Since $\i'$ is $R$-balanced, there is a unique group homomorphism $f:M
    \otimes_R N \xrightarrow{} D$ such that $\i'=f \circ \i$. Then $\i'=(f(m
    \otimes n))$. Since we also have that $D=\langle \i'(M \times N) \rangle$,
    we get that $f$ is onto. Since $\i'$ is $R$-balanced, there exists a unique
    group homomorphism $g:D \xrightarrow{} M \otimes_R N$ with $\i=g \circ \i'$.
    Then $m \otimes n=(g \circ f)(m \otimes n)$, and since every $m \otimes n$
    generates $M \otimes_R N$, we get $g \circ f=i_D$ (the identity on $D$), so
    that $f$ is 1--1.
\end{proof}

\begin{definition}
    Let $R$ and $S$ be rings with identity. An Abelian group $M$ is caled an
    \textbf{$(S,R)$-bimodule}  if $M$ is a left $S$-module, a right $R$-module,
    and $s(mr)=(sm)r$ for all $s \in S$, $r \in R$, and $m \in M$.
\end{definition}

\begin{example}\label{example_4.10}
    \begin{enumerate}
        \item[(1)] Let $S$ be a ring with identity and $R$ a subring of $S$ with
            $1_R=1_S$. Then $S$ is an $(S,R)$-bimodule by associativity in $S$.
            Now, if $f:R \xrightarrow{} S$ is a ring homomorphism with
            $f(1_R)=1_S$, then $S$ is a right $R$-module under the action
            $sr=sf(r)$.

        \item[(2)] Let $I$ be an ideal of a ring $R$. The quotient ring
            $\faktor{R}{I}$ is an $(\faktor{R}{I},R)$-bimodule.

        \item[(3)] Let $R$ be a commutative ring (not necessarily with
            identity), then any $R$-module $M$ is an $(R,R)$-bimodule, since
            $rm=mr$ for all $m \in M$ and $r \in R$. We call $M$ the
            \textbf{standard $R$-module structure} over $R$.

        \item[(4)] Let $R$ and $S$ be rings where $R \subseteq Z(S)$. Let $M$ be
            a left $S$-module, then $R$ is a commutative ring, and we have for
            all $s \in S$, $r \in R$, and $m \in M$ that
            $(sm)r=r(sm)=(sr)m=s(rm)=s(mr)$. So $M$ can be considered as a right
            $R$-module. This makes $M$ an $(S,R)$-bimodule under the right
            action of $R$.
    \end{enumerate}
\end{example}

\begin{definition}
    Let $R$ be a commutative ring with identity, and let $M$ and $N$, and $L$ be
    left $R$-modules. We call a map $\phi:M \times N \xrightarrow{} L$
    \textbf{$R$-bilinear} if it is $R$-linear in each factor. That is,
    \begin{align*}
        \phi(r_1m_1+r_2m_2,n)   &=  r_1\phi(m_1,n)+r_2\phi(m_2,n)   \\
        \phi(m,r_1n_1+r_2n_2)   &=  r_1\phi(m,n_1)+r_2\phi(m,n_2)   \\
    \end{align*}
    for all $r_1,r_2 \in R$, $m_1,m_2,m \in M$, and $n_1,n_2,n \in N$.
\end{definition}

\begin{lemma}\label{4.4.7}
    Let $R$ be a commutative ring, and let $M$ and $N$ be left $R$-modules. Let
    $M$ have the standard $R$-module structure. Then $M \otimes_R N$ is a left
    $R$-module with
    \begin{equation*}
        r(m \otimes n)=rm \otimes n=m \otimes (rn)
    \end{equation*}
    Moreover, the map $\i:(m,n) \xrightarrow{} m \otimes n$ is $R$-bilinear.
\end{lemma}
\begin{proof}
    Since $M$ and $N$ are left $R$-modules, and since $M$ has an
    $(R,R)$-bimodule structure, the tensor product $M \otimes_R N$ is a left
    $R$-module. Moreover, notice that $\i:(m,n) \xrightarrow{} m \otimes n$ is
    additive in each factor since $r(m \otimes n)=rm \otimes n=mr \otimes n=m
    \otimes rn$. This makes $\i$ a bilinear map.
\end{proof}
\begin{corollary}
    There exists a 1--1 correspondence of $R$-bilinear maps from $M \times N
    \xrightarrow{} L$ onto $R$-module homomorphisms from $M \otimes_R N
    \xrightarrow{} L$ given by the following diagram
    \[\begin{tikzcd}
        {M \times N} & {M \otimes_R N} \\
        & L
        \arrow["\iota", from=1-1, to=1-2]
        \arrow["\Phi", from=1-2, to=2-2]
        \arrow["\phi"', from=1-1, to=2-2]
    \end{tikzcd}\]
\end{corollary}
\begin{proof}
    If $\phi:M \times N \xrightarrow{} L$ is bilinear, then it is $R$-balanced.
    So the corresponding map $\Phi:M \otimes_R N \xrightarrow{} L$ is a group
    homomorphism. Notice now that
    \begin{equation*}
        \Phi((rm) \otimes n)=\phi(rm,n)=r\phi(m,n)=r\Phi(m \otimes n)
    \end{equation*}
    Since $\Phi$ is additive, this makes $\Phi$ a ring homomorphism. Now, if
    $\Phi$ is an $R$-module homomorphism with $\phi=\Phi \circ \i$, since $\i$
    is a bilenar map, it is easy to check that $\phi$ must also be bilinear.
\end{proof}

\begin{example}\label{example_4.11}
    \begin{enumerate}
        \item[(1)] In $M \otimes_R N$, $m \otimes 0=m \otimes (0+0)=m \otimes
            0+m \otimes 0$, so $m \otimes 0=0$.

        \item[(2)] Notice for any $a \in \faktor{\Z}{2\Z}$, $3a=a$, so for every
            $b \in \faktor{\Z}{3\Z}$, $a \otimes b=3a \otimes b=a \otimes 3b=a
            \otimes 0=0$. This shows that every simple tensor in
            $\faktor{\Z}{2\Z} \otimes_\Z \faktor{\Z}{3\Z}$ reduces to $0$. In
            particular, $1 \otimes 1=0$. Thus there are no nonzero bilinear maps
            mapping $\faktor{\Z}{2\Z} \otimes_\Z \faktor{\Z}{3\Z}$ onto any
            Abelian group.

        \item[(3)] Consider $\faktor{\Z}{2\Z} \otimes_\Z \faktor{\Z}{2\Z}$. This
            tensor product is generated by $0 \otimes 0=1 \otimes 0=0 \otimes
            1=0$, and $1 \otimes 1$. The map
            \begin{eqnarray*}
                \faktor{\Z}{2\Z} \otimes_\Z \faktor{\Z}{2\Z} & \xrightarrow{} &
                                             \faktor{\Z}{2\Z}   \\
                a \otimes b     & \xrightarrow{} & ab \mod{2}   \\
            \end{eqnarray*}
            is a nonzero and bilenear map. Moreover, notice that $2(1 \otimes
            1)=0$, so that $\ord{(1 \otimes 1)}=2$. That is
            \begin{equation*}
                \faktor{\Z}{2\Z} \otimes_\Z \faktor{\Z}{2\Z} \simeq \faktor{\Z}{2\Z}
            \end{equation*}

        \item[(4)] In general, whenever $m,n \in \Z^+$ have greatest common
            divisor $d$, then
            \begin{equation*}
                \faktor{\Z}{m\Z} \otimes_\Z \faktor{\Z}{n\Z} \simeq \faktor{\Z}{d\Z}
            \end{equation*}
            Moreover, observe that $a \otimes b=ab(1 \otimes 1)$, so that
            \begin{equation*}
                \faktor{\Z}{m\Z} \otimes_\Z \faktor{\Z}{n\Z}=\langle 1 \otimes 1 \rangle
            \end{equation*}
            making it cyclic. Moreover, $\ord{\langle 1 \otimes 1\rangle}|d$.

            Now, the map
            \begin{eqnarray*}
                \faktor{\Z}{m\Z} \times \faktor{\Z}{n\Z} & \xrightarrow{} &
                                             \faktor{\Z}{d\Z}   \\
                a \times b     & \xrightarrow{} & ab \mod{d}   \\
            \end{eqnarray*}
            is well defined since $d=(m,n)$. Moreover, $\phi$ is a $\Z$-bilinear
            map. Now, the induced map $\Phi:\faktor{\Z}{m\Z} \otimes_\Z
            \faktor{\Z}{m\Z} \xrightarrow{} \faktor{\Z}{d\Z}$ maps $1 \otimes 1
            \xrightarrow{} 1 \mod{d}$, so that $\faktor{\Z}{m\Z} \otimes_\Z
            \faktor{\Z}{m\Z}$ has order atleast $d$. Since $\faktor{\Z}{m\Z}
            \otimes_\Z \faktor{\Z}{n\Z}$ is cyclic and has order dividing $d$,
            by Lagrange's theorem its order is preciesly $d$. This proves the
            isomorphims.
    \end{enumerate}
\end{example}
