\section{The Path Integral}

\begin{definition}
    Let $a,b \in \R$ with  $a<b$. A complex valued function $\y:[a,b]
    \xrightarrow{} \C$ is said to be of \textbf{bounded variation} if there
    exists an $M>0$ such that for any partition $P=\{a=t_0<\dots<t_n=\}$ of
    $[a,b]$ the sum
    \begin{equation*}
        v(\y,P)=\sum_{k=0}^{n-1}{|\y(t_{k+1})-\y(t_k)|} \leq M
    \end{equation*}
    We call $v(\y,P)$ the \textbf{variation} of $\y$ on  $P$. We define the
     \textbf{total variation} of $\y$ on  $[a,b]$ to be
     \begin{equation*}
         V(\y)=\sup{\{v(\y,P)\}}
     \end{equation*}
     for every partition $P$ of  $[a,b]$.
\end{definition}

\begin{lemma}\label{4.1.1}
    Let $\y$ be a function of bounded variation. Then there exists an  $M>0$
    for which  $V(\y) \leq M$.
\end{lemma}

\begin{lemma}\label{4.1.2}
    Let $\y(t)=u(t)+iv(t)$ be a complex valued function. Then $\y$ is of bounded
    variation if, and only if $u$ and $v$ are of bounded variation.
\end{lemma}

\begin{lemma}\label{4.1.3}
    If $\y$ is a realvalued nondecreasing function, then  $\y$ is of bounded
    variation, and
    \begin{equation*}
        V(\y)=\y(b)-\y(a)
    \end{equation*}
\end{lemma}

\begin{lemma}\label{4.1.4}
    Let $\y:[a,b] \xrightarrow{} \C$ be a function of bounded variation. THen
    the following are true.
    \begin{enumerate}
        \item[(1)] If $P$ is a partition of  $[a,b]$, and $Q$ is a refinement of
             $P$, then  $v(\y,P) \leq v(\y,Q)$.

         \item[(2)] If $\s:[a,b] \xrightarrow{} \C$ is a function of bounded
             variation, and $\a,\b \in \C$, then  $\a\y+\b\s$ is of bounded
             variation and $V(\a\y+\b\s)=|\a|V(\y)+|\b|V(\s)$.
    \end{enumerate}
\end{lemma}

\begin{theorem}\label{4.1.5}
    Let $\y:[a,b] \xrightarrow{} \C$ be a piecewise $C^1$ function. Then $\y$ is
    of bounded variation and
    \begin{equation*}
        V(\y)=\int_a^b{|\y'(t)| \ dt}
    \end{equation*}
\end{theorem}
\begin{proof}
    Suppose without loss of generality that $\y$ is of class $C^1$ on all
    $[a,b]$. Let $P=\{a=t_0<\dots<t_n=b\}$ a partition of $[a,b]$. Then by
    definition, we have
    \begin{equation*}
        v(\y,P)=\sum_{k=0}^{n-1}{|\y(y_{k+1}\y(y_k))|}=
        \sum_{k=0}^{n-1}{\Big{|} \int_{t_k}^{t_{k+1}}{\y'(t) \ dt} \Big{|}} \leq
        \sum_{k=0}^{n-1}{\int_{t_k}^{t_{k+1}}{|\y'(t)| \ dt}}=
        \int_a^b{|\y'(t)| \ dt}
    \end{equation*}
    This gives $V(\y) \leq \int_a^b{|\y'|}$

    Now, since $\y'$ is continuous, it is uniformly continuous. Let  $\e>0$ and
    choose  $\d_1>0$ such that $|\y'(s)-\y'(t)|<\e$, whenever $0<|s-t|<\d_1$.
    Choose a $\d_2>0$ such that if $\|P\|=\max{\{t_{k+1}-t_k\}}<\d_2$, the
    \begin{equation*}
        \Big{|} \int_a^b{|\y(t)| \ dt}-
        \sum_{k=0}^{n-1}{\y'(\t_k)}|t_{k+1}-t_k| \Big{|}<\e
    \end{equation*}
    where $\t_k \in [t_k,t_{k+1}]$. Then we get
    \begin{equation*}
    \int_a^b{|\y'(t)| \ dt} \leq \e+\sum_{k=0}^{n-1}{|\y'(\t_k)|(t_{k+1}-t_k)}=
    \e+\sum_{k=0}^{n-1}{\Big{|} \int_a^b{|\y(\t_k)| \ dt} \Big{|}}
    \end{equation*}
    Then we get
    \begin{equation*}
        \int_a^b{|\y'(t)| \ dt} \leq
        \e+\sum_{k=0}^{n-1}{\Big{|} \int_{t_k}^{t_{k+1}}{\y'(\t_k)-\y'(t) \ dt}\Big{|}}+
        \sum_{k=0}^{n-1}{\Big{|} \int_a^b{\y'(t) \ dt} \Big{|}}
    \end{equation*}
    Now, take $\d=\min{\{\d_1,\d_2\}}$. If $\|P\|<\d$, then we get
    \begin{equation*}
        |\y'(\t_k)-\y'(t)|<\d \text{ for all } \t_k \in [t_k,t_{k+1}]
    \end{equation*}
    and
    \begin{equation*}
        \int_a^b{|\y'(t)| \ dt} \leq
        \e+\e(b-a)+\sum_{k=0}^{n-1}{|\y(t_{k+1}-\y(t_k))|}=\e+\e(b-a)+v(\y,p)
        \leq
        \e+\e(b-a)+V(\y)
    \end{equation*}
    Then, as $\e \xrightarrow{} 0$ from the positive side we observe that
    $\int_a^b{|\y'|} \leq V(\y)$ and equality is established.
\end{proof}

\begin{theorem}\label{4.1.6}
    Let $y:[a,b] \xrightarrow{} \C$ be a function of bounded variation, and
    suppose that $f:[a,b] \xrightarrow{} \C$ is a continuous complex valued
    function. Then there exists an $I \in \C$ such that for every $\e>0$, there
    is a $\d>0$ such that whenever  $P=\{a=t_0<\dots<t_n=b\}$ is a partition of
    $[a,b]$, with $\|P\|=\max{\{t_{k+1}-t_k\}}<\d$, then
    \begin{equation*}
        \Big{|} I-\sum_{k=0}^n{f(\t_k)(\y(t_{k+1}-\y(t_k)))} \Big{|}<\e
    \end{equation*}
    for any $\t_k \in [t_k,t_{k+1}]$.
\end{theorem}
