\section{Cyclotomic Fields and Cyclotomic Polynomials}
\label{section_8.7}

\begin{definition}
  Let $F$ be a field. We call the roots of the polynomial $x^n-1$ over
  $F$ the \textbf{n}-th roots of unity over $F$. If  $\xi$ is an
  $n$-th root of unity over $F$, we call the splitting field $F(\xi)$
  of $x^n-1$ over $F$ the \textbf{cyclotomic field} of $n$-th roots of
  unity over $F$.
\end{definition}

\begin{proposition}\label{proposition_8.7.1}
  Let $F$ be a field, and let $\xi$ and $\z$  $n$-th roots of unity
  over $F$, then $F(\xi) \simeq F(\z)$. That is: cyclotomic fields are
  unique up to isomorphism.
\end{proposition}
\begin{proof}
  $F(\xi)$ and $F(\z)$ are splitting fields for $x^n-1$ over $F$, so
  the result follows from the corollary to theorem
  \ref{theorem_8.4.4}.
\end{proof}

\begin{proposition}\label{proposition_8.7.2}
  Let $F$ be a field, and $\Xi_n(F)$ the set of all $n$-th roots of
  unity over $F$. Then $\Xi_n(F)$ is a cyclic group under the
  multiplication of the cyclotomic field over $F$. Indeed, it is a
  subgroup of the cyclotomic field over $F$.
\end{proposition}
\begin{proof}
  Take $\a,\b \in \Xi_n(F)$. Then $\a^n=1$ and $\b^n=1$. Then
  $(\a\b)^n=\a^n\b^n=1$. Moreover, $(\inv{\a})^n=\inv{(\a^n)}=1$.
  This makes $\Xi_n(F)$ a subgroup of the cyclotomic field $F(\xi)$,
  where $\xi$ is an $n$-th root of unity.

  Now, take $\xi$ an $n$-th root of unity over $F$. Then $\xi^n=1$.
  Consider $0 \leq k \leq n$, then $(\xi^k)^n=(\xi^n)^k=1$, so
  that $\xi^k$ is also an $n$-th root of unity. This makes $\Xi_n(F)$
  cyclic.
\end{proof}

\begin{definition}
  Let $F$ be a field, and $\Xi_n(F)$ the cyclic group of $n$-th roots
  of unity over $F$. We call a generator of $\Xi_n(F)$ a
  \textbf{primitive} $n$-th root of unity over $F$.
\end{definition}

\begin{proposition}\label{proposition_8.7.3}
  Let $F$ be a field. Then an $n$-th root of unity $\xi \in \Xi_n(F)$
  is primitive if, and only if its multiplicative order is coprime to
  $n$.
\end{proposition}
\begin{proof}
  This follows from results from cyclic groups; i.e. see lemma
  \ref{lemma_2.3.7}.
\end{proof}

\begin{proposition}\label{proposition_8.7.4}
  Let $F$ be a field. Then $\Xi_n(F) \simeq \faktor{\Z}{n\Z}$.
\end{proposition}
\begin{proof}
  The map $a \xrightarrow{} \xi^a$ defines the required isomorphism.
\end{proof}
\begin{corollary}
  There are exactly $\phi(n)$ primitive $n$-th roots of unity over
  $F$.
\end{corollary}
\begin{proof}
  Since $\xi^n \equiv \xi^{0 \mod{n}} \equiv 1$, we have every non identity
  power of $\xi$ has exponenct coprime to  $n$. That is there are $\phi(n)$
  such distinct powers of$\xi$.
\end{proof}
\begin{corollary}
  If $d|n$, then  $\Xi_d \leq \Xi_n$.
\end{corollary}
\begin{proof}
  Notice that if $d|n$, then $d=mn$ for some  $m \in \Z^+$. Then $\xi^d=1$
  implies  $(\xi^d)=\xi^{dm}=\xi^n=1$.
\end{proof}

\begin{definition}
  Let $F$ be a field. We define the \textbf{$n$-th cyclotomic
  polynomial} to be the polynomial $\Phi_n(x)$ over $F$,
  having as roots all $n$-primitive roots of unity over $F$.
\end{definition}

\begin{proposition}\label{proposition_8.7.5}
  The $n$-th cyclotomic polynomial  $\Phi_n(x)$ has degree
  $\deg{\Phi_n}=\phi(n)$, where $\phi$ is Euler's totient.
\end{proposition}
\begin{proof}
  Recall that $\ord{\Xi_n}=\phi(n)$, and since the elements of $\Xi_n$ are the
  roots of  $\Phi_n$, there are  $\phi(n)$ such roots. This puts
  $\deg{\Phi_n}=\phi(n)$.
\end{proof}

\begin{example}[Computing Cyclotomic Polynomials]\label{8.18}
  Recall that the polynomial $x^n-1$ has as roots precisely all  $n$-th roots
  of unity  $\xi$, that is $\xi^n=1$. If $x^n-1 \in F[x]$, $F$ a field, the
  the splitting field of  $x^n-1$ is  $F(\xi)$. Then we have
  \begin{equation*}
    x^n-1=\prod_{\xi \in \Xi_n(F)}{(x-\xi)}
  \end{equation*}
  Now, grouping those factors where $\xi^d=1$ for some  $d|n$, then we have
  \begin{equation*}
    x^n-1=\prod_{\xi \in \Xi_d}{(x-\xi)}\prod_{\xi \in \Xi_n}{(x-\xi)}=
    \prod{d|n}{\prod_{\xi \in \Xi_n}{(x-\xi)}}=\prod_{d|n}{\Phi_n(x)}
  \end{equation*}
  that is,
  \begin{equation*}
    x^n-1=\prod_{d|n}{\Phi_d(x)}
  \end{equation*}
  which gives a method for computing $\Phi_n$ recursively.

  We have  $\Phi_1(x)=x-1$ and $\Phi_2(x)=x+1$. Now,
  $\Phi_3(x)=\Phi_1(x)\Phi_3(x)=(x-1)\Phi_3(x)$, so that
  \begin{equation*}
    \Phi_n(x)=x^2+x+1
  \end{equation*}
  We have
  $\Phi_4(x)=\Phi_1(x)\Phi_2(x)\Phi_4(x)=(x-1)(x+1)\Phi_4(x)=(x^2-1)\Phi_4(x)$.
  So
  \begin{equation*}
    \Phi_4(x)=x^2+1
  \end{equation*}
  Similarly,
  \begin{align*}
    \Phi_5(x)   &=  x^4+x^3+x+1 \\
    \Phi_6(x)   &=  x^2-x+1 \\
    \Phi_7(x)   &=  x^6+x^5+x^4+x^3+x^2+x+1 \\
    \Phi_8(x)   &= x^4+1    \\
    \Phi_9(x)   &=  x^6+x^x+1   \\
    \Phi_{10}(x)   &=   x^4-x^3+x^2-x+1 \\
    \Phi_{11}(x)   &=   x^{10}+x^9+x^8+x^7+x^6+x^5+x^4+x^3+x^2+x+1  \\
    \Phi_{12}(x)   &=   x^4-x^2+1   \\
  \end{align*}
  Also observe that if $p$ is prime, then
  \begin{equation*}
    \Phi_p(x)=\sum_{i=0}^{p-1}{x^i}=x^{p-1}+x^p+\dots+x+1
  \end{equation*}
\end{example}

\begin{proposition}\label{proposition_8.7.6}
  $\Phi_n(x)$ is monic over $\Z$.
\end{proposition}
\begin{proof}
  Notice that since $x^n-1=\prod{\Phi_d(x)}$, is monic, then each $\Phi_d$
  must also be monic for all  $d|n$.

  Now, by induction on $n$, for  $n=1$, it is clear that  $x-1$ has
  coefficiencts in  $\Z$  (if $x^n-1 \in \Z[x]$ we are done, if not, just take
  $1_F \xrightarrow{} 1_\Z$, whre $F$ is the underlying field of  $x^n-1$).
  Now, suppose that $\Phi_d(x) \in \Z[x]$ for all $1 \leq d <n$, and  $d|n$.
  Then  $x^n-1=f(x)\Phi_n(x)$, where $f(x)=\prod{\Phi_d(x)}$ is monic over
  $\Z$. Moreover,  $f|x^n-1$, in the splitting field $\Q(\xi)$ (since we take
  $1_F \xrightarrow{} F_\Z$), where $\xi^n=1$. Then  $f|x^n-1$ over  $\Q$ by
  the division theorem, and by Gauss' proposition, $f|x^n-1$ in $\Z$. So $\Phi_n \in
  \Z[x]$.
\end{proof}
\begin{corollary}
  For any field $F$, and  $\xi$ a primitive $n$-th root of unity,
  $F(\xi) \simeq \Q(\xi)$.
\end{corollary}

\begin{theorem}\label{theorem_8.7.4}
  $\Phi_n(x)$ is irreducible over  $\Z$.
\end{theorem}
\begin{proof}
  Again, if $x^n-1 \in F[x]$ for some field $F$, take  $1_F \xrightarrow{}
  1_\Z$ so that $x^n-1 \in \Z[x]$. Suppose then that $\Phi_n(x)=f(x)g(x)$
  where $f$ and  $g$ are monic, and  $f$ is irreducible. Let $\xi^n=1$, a
  primitive $n$-th root, so that $\xi$ is a root of $f$. Then $f$ is the
  minimal polynomial for $\xi$ over  $\Q$. Now, let $p$ a prime such that $p
  \nmid n$. Then $\xi^p$ is a $n$-th root, of $f$ or $g$. If  $f(\xi^p)=0$,
  then for all $a$ with  $(a,n)=1$, we have $\xi^a$ is a root of $f$.
  Moreover,  $a=p_1 \dots p_k$ where each $p_i \nmid n$ is prime. That means
  tht  $\xi^{p_1}$, $(\xi^{p_1})^{p_2}, \dots, \xi^n$ are all roots of $f$
  making  $f=\Phi_n$ and we are done.

  Suppose then that $g(\xi^p)=0$. Then $\xi$ is root of  $g(x^p)$,
  and since $f$ is minimal, $f|g(x^p)$ in $\Z[x]$. Then we have
  $g(x^p)=f(x)h(x)$ for $f,h \in \Z[x]$. reducing mod $p$, we get
  $g(x^p) \equiv f(x)h(x) \mod{p}$ in $\F_p[x]$; but $g(x^p) \equiv (g(x))^p
  \mod{p}$. Since $\F_p[x]$ is a unique factorization domain, we get that $f
  \mod{p}$ and $g \mod{p}$ have a common factor. Then $\Phi_n(x) \equiv
  f(x)g(x) \mod{p}$ has a multiple root in $\F_p[x]$ ; implying that $x^n-1$
  has a multiple root, which is impossible; since $x^n-1$ has  $n$ distinct
  roots. Therefore $\xi^p$ is a root of  $f$.
\end{proof}
\begin{corollary}
  For any field $F$ and $\xi$ a primitive $n$-th root of unity,
  $[F(\xi):F]=\phi(n)$. In particular, $[\Q(\xi):\Q]=\phi(n)$.
\end{corollary}
\begin{proof}
  We have by above that $\Phi_n$ is the minimal polynomial for  $\xi$ over
  $\Q$.
\end{proof}

\begin{example}\label{8.19}
  Let $\xi^8=1$ an  $8$-th root of unity. Then  $[\Q(\xi):\Q]=4$ and $\Q(\xi)$
  has minimal polynomial $\Phi_8(x)=x^4+1$. Moreover, $\Q(\xi)$ contains a
  primitive $4$-th root of unity $i^4=1$ (over $\C$,  $i^2=-1$). So that
  $\Q(i) \subseteq \Q(\xi)$. We als get that $\xi+\xi^7=\sqrt[]{2}$ (since
  $\xi=e^{\frac{2i\pi}{8}}$ over $\C$), and $\Q(\xi)=\Q(i,\sqrt[]{2})$.
\end{example}
