\section{Definitions and Equivalences of Boolean Functions}\label{section_1.1}

\begin{definition}
  Let $p,m, n \in \Z^+$, where $p$ is prime. By an \textbf{$(n,m,p)$-function},
  we mean a function $F:\F_p^n \xrightarrow{} \F_p^m$. If $p=2$, we call $F$ an
  \textbf{$(n,m)$-Boolean function}, or an \textbf{$S$-box}. If
  $F(x_1, \dots, x_n)=(f_1(x_1, \dots, x_n), \dots, f_m(x_1, \dots, x_n))$, we
  call each $f_i$ the \textbf{$i$-th coordinate} function of $F$.
\end{definition}

\begin{lemma}\label{lemma_1.1.1}
  If $F$ is an  $(n,m,p)$-function, tthen $F$ can be represented uniquely as a
  univariate polynomial over  $\F_{p^n}$ of the form:
  \begin{equation}\label{equation_1.1}
    F(x)=\sum_{i=0}^{p^n-1}{c_ix^i},  \text{ where } c_i \in \F_{p^n}
    \text{ for all } 0 \leq i \leq p^n-1
  \end{equation}
\end{lemma}
\begin{proof}
  Recall that $\F_p^r \simeq \F_{p^r}$ for any integer $r \in \Z^+$. Then any
  $(n,m,p)$-function can be considered as a function taking elements of
  $\F_{p^n}$ to elements of $\F_{p^n}$, so that there is a unique polynomial
  representation of $F$ of the form in equation \ref{equation_1.1}.
\end{proof}

\begin{defintion}
  Let $k$ be an integer where $0 \leq k \leq p^n-1$. We define the
  \textbf{$p$-weight} of $k$ to be
  \begin{equation}\label{equation_1.2}
    w_p(k)=\sum_{s=0}^{n-1}{k_s} \text{ for all } 0 \leq k_s \leq p-1
  \end{equation}
  in the $p$-ary expansion  $\sum_{s=0}^{n-1}{p^sk_s}$. We define the
  \textbf{algebraic degree} of an $(n,m,p)$-function $F$ to be
  \begin{equation}\label{equation_2.3}
    d^\circ(F)=\max{\{w_p(i) : c_i \neq 0 \text{ and } 0 \leq i \leq p^n-1\}}
  \end{equation}
  where $F$ is represented by equation \ref{equation_1.1}.
\end{defintion}

\begin{definition}
  We call an $(n,n,p)$-function $F$ is called \textbf{linear} if $F$ can be
  represented as a polynomial in $\F_{p^n}[x]$ of degree at most $n-1$. We call
   $F$ \textbf{affine} if it is the sum of linear functions and a constant.
\end{definition}

\begin{definition}
  We call an $(n,n,p)$-function \textbf{DEmbowski-Ostrom polynomial}, or
  \textbf{DO-polynomial} if
  \begin{equation}\label{equation_1.4}
    F(x)=\sum_{0 \leq k,j<n}{a_{kj}x^{p^k+p^j}} \text{ where } a_{ij} \in
    \F_{p^n}
  \end{equation}
  We call $F$ if it is the sum of  DO-polynomials and a constant.
\end{definition}

\begin{defintion}
  Let $m$ be a positive divisor of $n$. We define the trace map
  $\tr_n^m:\F_{p^n} \xrightarrow{} \F_{p^m}$ defined by:
  \begin{equation*}
    \tr_n^m(x)=x+x^{p^m}+ \dots +x^{p^{\frac{nm}{m-1}}}
  \end{equation*}
  When $m=1$, we write $\tr_n(x)$, when $n=2$, we write $\tr^m(x)$, and when
  $n=2$ and $m=1$ we write $\tr(x)$.
\end{defintion}

\begin{lemma}\label{lemma_1.1.2}
  Let  $F$ be an $(n,m,p)$-function, then there exists an $(n,n,p)$-function $G$
  for which $F(x)=\tr_n^m(G(x))$. Moreover, any $(n,1,p)$-functioin $F$ can be
  written in a nonunique way as $F(x)=\tr_n(G(x))$, where $G \in \F_{2^n}[x]$,
  and the univariate representation can be written as:
  \begin{equation}\label{equation_1.5}
    F(x)=\sum_{j \in \Gamma_n}{\tr_{o(j)}(a_jx^j)} \text{ where } a_j \in
    \F_p^{o(j)}
  \end{equation}
  where $\Gamma_n$ is the choice set obtained by choosing one representative
  from each cyclotomic coset of $p \mod{p^n-1}$, and $o(j)$ is the size of the
  cyclotomic coset containing both $j$ and  $a_j$.
\end{lemma}

\begin{lemma}\label{1.1.3}
  Let $n=2k$, and $F$ be an $(n,m,p)$-function. Then there exists a unique
  bivariate polynomial over $\F_{p^k}$ representing $F$ of the form:
  \begin{equation}\label{equation_1.6}
    F(x)=\sum_{0 \leq i,j, \leq p^k-1}{a_ijx^iy^j} \text{ where } a_{ij} \in
    \F_{p^k}
  \end{equation}
  Moreover
  \begin{equation*}
    d^\circ(F)=\max{\{w_p(i)+w_p(j) : (i,j)|a_{ij} \neq 0\}}
  \end{equation*}
  and
  \begin{equation*}
    F(x,y)=\tr_k(P(x,y)) \text{ where } P \in \F_{p^k}[x,y]
  \end{equation*}
\end{lemma}

\begin{definition}
  We define the \textbf{algebraic normal form} of an $(n,m,p)$-function $F$ to
  be the representation:
  \begin{equation}\label{equation_1.7}
    F(x)=\sum_{u \in \F_p^n}{a_u\prod_{i=1}^nx_i^{u_i}}, \text{ where }
    a_u \in \F_p^m
  \end{equation}
  We define the \textbf{minimum degree} of $F$ to be
  $d^\circ=\min{d^\circ(F)}$, the minimum degree of all non-zero linear
  combinations of the coordinate functions of $F$.
\end{definition}

\begin{definition}
  We call an $(n,m,p)$-function \textbf{balanced} if every element of $\F_p^m$
  has the same number $p^{n-m}$ of pre-images.
\end{definition}

\begin{defition}
  Let $\d \in \Z^+$. We call an $(n,m,p)$-function \textbf{differentially
  $\d$-uniform} if for every non-zero $a \in \F_p^n$, and every $b \in \F_p^m$,
  the differnce equation:
  \begin{equation}\label{equation_1.8}
    F(x+a)-F(x)=b
  \end{equation}
  has at most $\d$-solutions. We call $F$ \textbf{perfect nonlinear (PN)} if it
  is differentially $p^{n-m}$-uniform. We define the \textbf{derivatve} of $F$
  \textbf{in the direction} of $a$ to be the function
  \begin{equation*}
    D_a{F(x)}=F(x+a)-F(x)
  \end{equation*}
\end{defition}

\begin{lemma}\label{lemma_1.1.4}
  An $(n,m,p)$-function is perfect nonlinear if, and only if $D_a{F}$ is
  balanced for every nonzero $a \in \F_p^n$.
\end{lemma}
\begin{corollary}
  If $n=m$, then $F$ is perfect nonlinear if, and only if $D_a{F}$ is a
  permutation.
\end{corollary}

\begin{definition}
  We call perfect nonlinear $(n,n,p)$-functions \textbf{planar} functions.
\end{definition}

\begin{example}
  There are no perfect nonlinear functions on $\F_2^n$, for any $n \in \Z^+$. Let
  $F$ be an $(n,m)$-Boolean function, and let $x_0$ be a solution to the equation
  $D_a{F(x)}=b$, where $a \neq 0$. Then $x_0+a$ is also a solution, since
  $D_a{F(x_0+a)}=F((x_0+a)+a)-F(x_0+a)=F(x_0+a)-F(x_0)=b$. So $F$ has at least
  $2$ solutions in $\F_2^n$. However, for  $F$ to be PN, $F$ must be
  differentially $1$-uniform. By the corollary to lemma \ref{lemma_1.1.4}, there
  are also no planar functions in $\F_2^n$.
\end{example}

\begin{definition}
  We call a Boolean function $F:F_2^n \xrightarrow{} \F_2^n$ \textbf{almost
  perfect nonlinear (APN)} if $D_a{F}$ is a 2--1 mapping; that is for every
  $a,b \in \F_2^n$, with $a \neq 0$, we have the equation
  \begin{equation*}
    F(x+a)-F(x)=b
  \end{equation*}
  has either exactly $0$, or  $2$ solutions.
\end{definition}

\begin{definition}
  We define the following relations on $(n,m,p)$-functions:
  \begin{enumerate}
    \item[(1)] We call $F$ and $G$ \textbf{affine equivalent} if there exist
      affine permutations $A_1:\F_p^m \xrightarrow{} \F_p^m$ and $A_2:\F_p^n
      \xrightarrow{} \F_p^n$ for which:
      \begin{equation*}
        G=A_1 \circ F \circ A_2
      \end{equation*}
      We call $F$ and $G$ \textbf{linear} equivalent if $A_1$ and $A_2$ are
      linear permutations.

    \item[(2)] We call $F$ and $G$ \textbf{extended affine equivalent
      (EA-equivalent)} if thre exist affine permutations $A_1:\F_p^m
      \xrightarrow{} \F_p^m$ and $A_2:\F_p^n \xrightarrow{} \F_p^n$, and an
      affine function $A:\F_p^n \xrightarrow{} \F_p^n$ for which:
      \begin{equation*}
        G=A_1 \circ A \circ A_2+A
      \end{equation*}

    \item[(3)] We call $F$ and $G$ \textbf{Carlet-Charpin-Zimoviev equivalent
      (CCZ-equivalent)} if for any affine permutation $\Lc$ of  $\F_p^n \times
      \F_p^m$, the image of the graph of $F$ is the graph of $G$; that is:
      \begin{equation*}
        \Lc(\F_p^n \times F(\F_p^n))=\F_p^n \times G(\F_p^n)
      \end{equation*}
  \end{enumerate}
\end{definition}

\begin{proposition}\label{proposition_1.1.5}
  The following are true for all $(n,m,p)$-functions:
  \begin{enumerate}
    \item[(1)] Affine equivalence is an equivalence relation.

    \item[(2)] EA-equivalence is an equivalence relation.

    \item[(3)] CCZ-equivalence is an equivalence relation.
  \end{enumerate}
\end{proposition}
\begin{corollary}
  Affine equivalence implies linear equivalence.
\end{corollary}
\begin{corollary}
  EA-equivalence implies affine equivalence. Moreover, the algebraic degree of a
  non-affine function is preserved under EA-equivalence.
\end{corollary}
\begin{corollary}
  CCZ-equivalence implies EA-equivalence. Moreover, every permutation is
  CCZ-equivalent to its inverse.
\end{corollary}

\begin{proposition}\label{proposition_1.1.6}
  Let $F$ and  $G$ be $(n,m,p)$-functions. If $\inv{F}$ exists, then $G$ is
  either EA-equivalent to $F$ or to $\inv{F}$ if, and only if there exists a
  linear permutation $\Lc=(L_1,L_2)$ on $\F_2^n \times F_2^m$ for which
  \begin{equation*}
    \Lc(\F_p^n \times F(\F_p^m))=\F_p^m \times G(\F_p^m)
  \end{equation*}
  and $L_1$ depends only on one variable.
\end{proposition}
