%----------------------------------------------------------------------------------------
%	SECTION 1.1
%----------------------------------------------------------------------------------------

\section{$\sigma$-algebras of Lebesuge Measurable Sets}

\begin{definition}
    We say a set $E$ is  \textbf{measurable} if for any $A \subseteq \R$, then:
    \begin{equation*}
        m^*(A)=m^*(A \cap E)+m*(A \cap (\com{A}{E}))
    \end{equation*}
\end{definition}

\begin{lemma}\label{8.2.1}
    A set $E$ is measurable if, and only if  $m^*(A) \geq m^*(A \cap E) +m^*(A
    \cap (\com{\R}{E}))$.
\end{lemma}

\begin{lemma}\label{8.2.2}
    Any set of outer measure $0$ is measurable.
\end{lemma}
\begin{proof}
    Let $E$ be a set of measure  $0$, and let  $A \subseteq \R$. Notice that  $A
    \cap E \subseteq E$ and  $A \cap (\com{\R}{E}) \subseteq A$. Then by
    monotonicity, $m^*(A \cap E) \leq m^*(E)=0$ and $m^*(A \cap (\com{\R}{E}))
    \leq m^*(A)$. Then we get that $m^*(A) \geq m^*(A \cap E)+m^*(A \cap
    (\com{\R}{E}))$.
\end{proof}
\begin{corollary}
    Countable sets are measurable.
\end{corollary}

\begin{lemma}\label{8.2.3}
    Unions of a finite collection of measurable sets is measurable.
\end{lemma}
\begin{proof}
    By induction, for $n=2$, let  $E_1$ and $E_2$ be measurable. Then for any $A
    \subseteq \R$, we get  $m^*(A)=m^*(A \cap E_1)+m^*(A \cap (\com{\R}{E_1}))=
    m^*(A \cap (\com{\R}{E_1}) \cap E_2)+m^*(A \cap (\com{\R}{E_1}) \cap
    (\com{\R}{E_2}))$. Notice that $A \cap (\com{\R}{E_1}) \cap
    (\com{\R}{E_2})=A \cap \com{{\R}}{(E_1 \cup E_2)}$, and that $(A \cap E_1)
    \cup (A \cap \com{\R}{E_1} \cap E_2)=A \cap (E_1 \cup E_2)$. Then we get
    that
    \begin{equation*}
        m^*(A) \geq m^*(A \cap (E_1 \cup E_2))+m^*(A \cap \com{\R}{(E_1 \cup E_2)})
    \end{equation*}
    This makes $E_1 \cup E_2$ measurable.

    Now, suppose that for $n \geq 2$ the collection $\{E_i\}_{i=1}^n$ of
    measurable set has measurable union. Consider now the collection $\{E_i\}
    \cup E_{n+1}$, where $E_{n+1}$ is also measurable. By hypothesis, we have :
    \begin{equation*}
        m^*(A) \geq m^*(A \cap \bigicup_{i=1}^{n+1}{E_i}) + m^*(A \cap
        \com{\R}{(\bigicup_{i=1}^{n+1}{E_i})})
    \end{equation*}
    therefore the collection $\{E_i\}_{i=1}^{n+1}$ of measurable sets has
    measurable union.
\end{proof}

\begin{lemma}\label{8.2.4}
    Let $\{E_i\}_{i=1}^{n}$ be a collection of mutually disjoint measurable
    sets. Then for any $A \subseteq \R$:
    \begin{equation*}
        m^*(A \cap \bigcup_{i=1}^n{E_i})=\sum_{i=1}^n{m^*(A \cap E_i)}
    \end{equation*}
\end{lemma}
\begin{proof}
    By induction on $n$, we have  $m^*(A \cap E_1)=\sum_{i=1}{1}{m^*(A \cap
    E)i)}$. Now, suppose that for $n \geq 1$, that
    \begin{equation*}
        m^*(A \cap \bigcup_{i=1}^n{E_i})=\sum_{i=1}^n{m^*(A \cap E_i)}
    \end{equation*}
    consider then the collection $\{E_i\} \cup E_{n+1}$ where $E_{n+1}$ is
    measurable and mutually disjoint from all $E_i$ for  $1 \leq i \leq n$.
    Notice that $(A \cap \bigcup{E_i}) \cap E_{n+1}=A \cap E_{n+1}$ and that $A
    \cap \bigcup_{i=1}^{n+1}{E_i} \cap \com{\R}{E_{n+1}}=A \cap \bigcup{E_i}$.
    Then by hypothesis, we have:
    \begin{equation*}
        m^*(A \cap \bigcup_{i=1}^{n+1}{E_i})=m^*(A \cap E_{n+1})+m^*(A \cap
    \bigcup_{i=1}^n{E_i})=m^*(A \cap E_{n+1})+\sum{m^*(A \cap
    E_i)}=\sum_{i=1}^{n+1}{m^*(A \cap E_i)}
    \end{equation*}
\end{proof}

\begin{definition}
    A collection of subsets of $\R$ is called an  \textbf{algebra} if it
    contains $\R$, and it is closed under complements and finite unions.
\end{definition}

\begin{theorem}\label{8.2.5}
    The collection of all measurable sets in $\R$ forms an algebra.
\end{theorem}
\begin{proof}
    By lemma \ref{8.1.3}, we have that this collection is closed under finite
    unions. Also observe that if $E$ is measruable, then  $m^*(A)=m^*(A \cap
    E)+m^*(A \cap \com{\R}{E})$ implies that $m^*(A)=m^*(A \cap
    \com{\R}{E})+m^*(A \cap \com{\R}{(\com{\R}{E})})$. This makes $\com{\R}{E}$
    measurable. Lastly we see that $m^*(A)=m^*(A \cap \R)+m^*(A \cap
    \com{\R}{\R})=m^*(A) + m^*(\emptyset)$, so that $\R$ is measurable.
    Therefore the collection of all measurable sets in  $\R$ is measurable.
\end{proof}
\begin{corollary}
    The collection of all measruable sets in $\R$ is closde under finite
    intersections.
\end{corollary}

\begin{lemma}\label{8.2.6}
    The union of a countable collection of measurable sets is measurable.
\end{lemma}
\begin{proof}
    Let $E=\bigcup{E_n}$ where $\{E_n\}_{n \in \Z^+}$ is a countable collection
    of measurable sets. Let $A \subseteq \R$, and for  $n \in \Z^+$, define
    $F_n=\bigcup_{k=1}^n{E_k}$. By lemma \ref{8.2.3}, $F_n$ is measurable for
    all  $n$ and  $\com{\R}{E} \subseteq \com{\R}{F_n}$. Then we have taht
    $m^*(A \cap F_n)+m^*(A \cap \com{\R}{F_n}) \geq m^*(A \cap F_n)+m^*(A \cap
    \com{\R}{E})$. By lemma \ref{8.2.4}, we get that $m^*(A \cap
    F_n)=\sum_{k=1}^n{E_k}$, thus we get that $m^*(A) \geq \sum{E_k}+m^*(A \cap
    \com{\R}{E})$. By countable subadditivity, this gives us $m^*(A) \geq m^*(A
    \cap E)+m^*(A \cap \com{\R}{E})$. Therefore $E$ is measurable.
\end{proof}

\begin{lemma}\label{8.2.7}
    Every real interval is measurable.
\end{lemma}
\begin{proof}
    Since the collection of all measurable sets form an algebra, in fact they
    form a $\sigma$-algebra, it suffices to show that  $(a,\infty)$ and
    $(-\infty,a)$ are measurable for all $a \in \R$.

    Let  $A \subseteq \R$, and suppose that  $a \notin A$ without loss of
    generality. Consider any countable collection  $\{I_n\}$ of open and bounded
    intervals covering $A$; and define  $I'_n=I_n \cap (-\infty,a)$ and
    $I''_n=I_n \cap (a,\infty)$. Now, $I'_n$ and  $I''_n$ are open intervals
    with $l(I_n)=l(I'_n)+l(I''_n)$ for all $n$. Since  $\{I_n\}$ is countable,
    so are the collections $\{I'_n\}$ and $\{I''_n\}$. They also cover subsets
    $A_1,A_2 \subseteq A$. Then $m^*(A_1)=\sum{l(I'_n)}$ and
    $m^*(A_2)=\sum{l(I''_n)}$, so tht $m^*(A_1)+m^*(A_2) \leq
    \sum{(l(I'_n)+l(I''_n))} \leq \sum{l(I_n)}$. This makes $(a,\infty)$ and $
    (-\infty,a)$ measurable.
\end{proof}
\begin{corollary}
    Any open set in $\R$ is measurable.
\end{corollary}

\begin{theorem}\label{8.2.8}
    The collection of al measurable sets containes the Borel algebra.
\end{theorem}

\begin{lemma}\label{8.2.9}
    Measurable sets are translation invariant.
\end{lemma}
\begin{proof}
    Let $E$ be a measurable set and let  $A \subseteq \R$ and  $y \in \R$. Then
     $m^*(A)=m^*(\com{A}{y})=m^*((\com{A}{y}) \cap E)+m^*((\com{A}{y}) \cap
     \com{\R}{E})=m^*(A \cap E+y)+m^*(A \cap \com{\R}{(E+y)})$.
\end{proof}
